---
title: "FML_Assignment_5"
output: html_document
date: "2024-04-04"
---

```{r}
setwd("/Users/meghana/Downloads")
data = read.csv("Cereals .csv")
```
```{r}
head(data)
```
```{r}
#Installing necessary packages
#install.packages("cluster")
#install.packages("caret")
#install.packages("dendextend")
#install.packages("knitr")
#install.packages("factoextra")
#install.packages("readr")
```

```{r}
library("cluster")
library("caret")
library("dendextend")
library("knitr")
library("factoextra")
library("readr")
```

```{r}
data <- data.frame(data[,4:16])
#Removing Null values
data = na.omit(data)
```

```{r}
#Normalizing data
Cereals_normalise <- scale(data)
```

```{r}
#Use the normalized data to do hierarchical clustering using the Euclidean Dist technique.
Dist <- dist(Cereals_normalise, method = "euclidean")
H_clust <- hclust(Dist, method = "complete")
```

```{r}
#the dendogram plotting
plot(H_clust, cex = 0.7, hang = -1)
```
```{r}
#Clustering with single linkage, full linkage, and the Agnes function,average linkage and Ward.
single_Hclust <- agnes(Cereals_normalise, method = "single")
complete_Hclust <- agnes(Cereals_normalise, method = "complete")
average_Hclust <- agnes(Cereals_normalise, method = "average")
ward_Hclust <- agnes(Cereals_normalise, method = "ward")
```

```{r}
#Choosing the most efficient course of action
print(single_Hclust$ac)
print(complete_Hclust$ac)
print(average_Hclust$ac)
print(ward_Hclust$ac)
#The ward strategy is the most successful one, as shown by its value of 0.9046042, which is evident given the facts provided.
```

### 2- Choosing the clusters:
```{r}
pltree(ward_Hclust, cex = 0.5, hang = -1, main = "Dendrogram of agnes (Using Ward)")
rect.hclust(ward_Hclust, k = 5, border = 2:7)
S_Group <- cutree(ward_Hclust, k=5)
D_frame_2 <- as.data.frame(cbind(Cereals_normalise,S_Group))
```

```{r}
#determining the stability and structure of the clusters. 
fviz_cluster(list(data = D_frame_2, cluster = S_Group))
```
```{r}
#Building Partitions
set.seed(123)
partition_1<- data[1:50,]
partition_2 <- data[51:74,]
```

```{r}
#Performing Hierarchical Clustering while considering k = 5.
single_sb <- agnes(scale(partition_1), method = "single")
complete_sb <- agnes(scale(partition_1), method = "complete")
average_sb <- agnes(scale(partition_1), method = "average")
ward_sb <- agnes(scale(partition_1), method = "ward")
cbind(single=single_sb$ac , complete=complete_sb$ac , average= average_sb$ac , ward= ward_sb$ac)
pltree(ward_sb, cex = 0.6, hang = -1, main = "Dendogram of Agnes with Partitioned Data (Using Ward)")
rect.hclust(ward_sb, k = 5, border = 2:7)
cut_2 <- cutree(ward_sb, k = 5)
```

```{r}
#the centroids are calculated.
Sb_result <- as.data.frame(cbind(partition_1, cut_2))
Sb_result[Sb_result$cut_2==1,]
one_centroid <- colMeans(Sb_result[Sb_result$cut_2==1,])
Sb_result[Sb_result$cut_2==2,]
two_centroid <- colMeans(Sb_result[Sb_result$cut_2==2,])
Sb_result[Sb_result$cut_2==3,]
three_centroid <- colMeans(Sb_result[Sb_result$cut_2==3,])
Sb_result[Sb_result$cut_2==4,]
four_centroid <- colMeans(Sb_result[Sb_result$cut_2==4,])
centroids <- rbind(one_centroid, two_centroid, three_centroid, four_centroid)
x2 <- as.data.frame(rbind(centroids[,-14], partition_2))
```
```{r}
#figuring out the Dist.
Dist_1 <- get_dist(x2)
Matrix_1 <- as.matrix(Dist_1)
dataframe1 <- data.frame(data=seq(1,nrow(partition_2),1), Clusters = rep(0,nrow(partition_2)))
for(i in 1:nrow(partition_2)) 
  {dataframe1[i,2] <- which.min(Matrix_1[i+4, 1:4])}
dataframe1
cbind(D_frame_2$S_Group[51:74], dataframe1$Clusters)
table(D_frame_2$S_Group[51:74] == dataframe1$Clusters)
```

### 3- The elementary public schools would like to choose a set of Cereals to include in their daily cafeterias. Every day a different cereal is offered, but all Cereals should support a healthy diet. For this goal, you are requested to find a cluster of “healthy Cereals''

```{r}
#Clustering Healthy Cereals.
Healthy_Cereals <- data
Healthy_Cereals_RD <- na.omit(Healthy_Cereals)
clust <- cbind(Healthy_Cereals_RD, S_Group)
clust[clust$S_Group==1,]
clust[clust$S_Group==2,]
clust[clust$S_Group==3,]
clust[clust$S_Group==4,]
```

```{r}
#Mean ratings are used to select the best cluster.
mean(clust[clust$S_Group==1,"rating"])
mean(clust[clust$S_Group==2,"rating"])
mean(clust[clust$S_Group==3,"rating"])
mean(clust[clust$S_Group==4,"rating"])
```
#### Cluster 1 may be chosen based on the data mentioned above because it is the highest.
#### Therefore, Group 1 may be considered of as the cluster for a healthy diet.
