---
title: "FML_Assignment_4"
output:
  pdf_document: default
  html_document: default
date: "2024-03-14"
---

# Cluster Analysis of Pharmaceutical Firms
## Introduction
### In this analysis, we perform cluster analysis on a dataset containing information about pharmaceutical firms. We focus on using numerical variables (1 to 9) to cluster the 21 firms. Various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, are justified.

```{r}
#Importing Required Packages
library(readr)
#Importing Data Set
data <- read_csv("/Users/meghana/Downloads/Pharmaceuticals.csv")
```

 

### Load necessary libraries
```{r}
library("ggplot2")
library("factoextra")
library("flexclust")
library("cluster")
library("tidyverse")
library("cluster")
```
#### Question(A) Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on.
```{r}
# Removing null values in data (data cleaning)
Pharma_data =  na.omit(data)
Pharma_data

row.names <- Pharma_data[,1]
pharma_data1 <- Pharma_data[,3:11]#numerical variable from 3 to 11
head(pharma_data1)
pharma_data2 <- scale(pharma_data1)
head(pharma_data2)
#Determination of Number of Clusters
#We determine the optimal number of clusters using different methods such as the Elbow Method, Silhouette Method, and Gap Stat Method.
fviz_nbclust(pharma_data2, kmeans, method = "wss") +labs(subtitle = "Elbow Method")
fviz_nbclust(pharma_data2, kmeans, method = "silhouette") + labs(subtitle = "Silhouette Method")
fviz_nbclust(pharma_data2, kmeans, method = "gap_stat") + labs(subtitle = "Gap Stat Method")
```

```{r}
set.seed(64060)
k_5 <- kmeans(pharma_data2, centers = 5, nstart = 25)
k_5 $centers
fviz_cluster(k_5, data = pharma_data2)
k_5
```


```{r}
distance <- dist(pharma_data2, method = "euclidian")
#fvi_dist(distance)
```


```{r}
FITT <- kmeans(pharma_data2,5)
aggregate(pharma_data2,by = list(FITT$cluster), FUN = mean)
pharma_data3 <- data.frame(pharma_data2,FITT$cluster)
pharma_data3
clusplot(pharma_data2,FITT$cluster, color = TRUE, shade = TRUE, 
 labels = 2,
 lines = 0)
```

#### Question(B) Interpret the clusters with respect to the numerical variables used in forming the clusters.

```{r}
aggregate(pharma_data2, by = list(FITT$cluster), FUN = mean)
Pharmacy <- data.frame(pharma_data2,k_5$cluster)
Pharmacy

# CLuster 1: JNJ, MRK, GSK, PFE (lowest beta/PE ratio and highest market cap)
#Cluster 2: AHM, WPI, AVE (lowest PE/Asset Turnover Ratio and highest revenue growth)
# Cluster 3: CHTT, IVX, MRX, ELN (lowest Net Profit Margin, PE ratio, and Marke#Cluster, and highest beta, leverage, and asset turnover ratio)
#Cluster 4: AGN, BAY, PHA (lowest leverage/asset turnover and highest PE ratio)
#BT, WYE, AZN, SGP, BMY, NVS, LLY are in Cluster 5 (highest net profit margin and #lowest leverage).
```



#### Question(C) Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)
```{r}
Pharma <- data[12:14] %>% mutate(Clusters=k_5$cluster)
ggplot(Pharma, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters')
ggplot(Pharma, mapping = aes(factor(Clusters),fill = Location))+geom_bar(position = 'dodge')+labs(x ='Clusters')
ggplot(Pharma, mapping = aes(factor(Clusters),fill = Exchange))+geom_bar(position = 'dodge')+labs(x ='Clusters')
```


#### Interpretation : 

#### The clusters from the graphs above show a slight pattern that we can observe #### The businesses in cluster 1 are evenly distributed across the AMEX, NASDAQ, and NYSE, but it has distinct Hold and Moderate Buy medians and a different count from the US and Germany.
#### The medians for holds and moderate purchases are distributed similarly in Cluster 2.
#### The NYSE lists stocks from both the US and Canada.
#### Cluster 3 differs from Cluster in count, but its Moderate Buy and Sell medians are comparable.
#### The NYSE lists France, Ireland, and the US. 
#### Hold, Moderate Buy, Moderate Sell, and Strong Buy options are available in Cluster 4.
#### The hold's median is the highest. They are listed on the NYSE and are citizens of the United States, the United Kingdom, and Switzerland.
#### Cluster 5 is spread across 
#### countries, including the US and the UK, and is listed on the NYSE. It also has the same hold and median purchase values.

#### Question(D) Provide an appropriate name for each cluster using any or all of the variables in the dataset.

```{r}
#Naming clusters
#After performing cluster analysis on the pharmaceutical firms dataset,Assigning  descriptive names to each cluster based on their characteristics.

#Cluster 1 :- Buy Cluster
#Cluster 2 :- Sceptical Cluster
#Cluster 3 :- Moderate Buy Cluster
#Cluster 4 :- Hold Cluster
#Cluster 5 :- High Hold Cluster
```


